---
title: "Getting started with pipelineR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with pipelineR}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

**pipelineR** connects three pieces:

1. **PostgreSQL** — symbols (`sp500.info`) and your schema (`data_sp500`, `pipeline_logs`).
2. **Yahoo Finance** (via `tidyquant::tq_get`) — daily OHLCV for those symbols.
3. **R** — batching, reshaping to long format, deduplicated inserts, and logging.

The usual entry point is `start_pipeline()`, which runs the full workflow and disconnects when finished.

## Configuration

Create or edit `~/.Renviron` (or project-level `.Renviron`) with:

| Variable | Role |
|----------|------|
| `PG_DB` | Database name |
| `PG_HOST` | Host |
| `PG_USER` | User |
| `PG_PASSWORD` | Password |
| `PG_SCHEMA` | Schema containing `data_sp500` and `pipeline_logs` |
| `user_login` | Your id (stored in `pipeline_logs`) |

`connect_db()` checks that `PG_DB`, `PG_HOST`, `PG_USER`, `PG_PASSWORD`, and `PG_SCHEMA` are non-empty before opening a connection.

## Run the pipeline

```{r eval = FALSE}
library(pipelineR)

start_pipeline(
  from = Sys.Date() - 10,
  to = Sys.Date(),
  batch_size = 25
)
```

Arguments:

- **`from`**, **`to`** — date range passed to Yahoo Finance (defaults: last 7 days through today).
- **`batch_size`** — number of symbols per API call (default `25`); smaller batches can reduce rate-limit issues.

## Pipeline steps (reference)

If you build a custom workflow, the logical order is:

1. `connect_db()` — DB connection  
2. `fetch_symbols(con)` — tickers from `sp500.info`  
3. `split_batch(symbols, batch_size)` — list of batches  
4. For each batch: `yahoo_query_data(batch, from, to)` then `format_data()` then `insert_new_data(con, ...)`  
5. `build_summary_table()` / `log_summary()` — in-memory log rows  
6. `push_summary_table(con, summary_table)` — append to `pipeline_logs`  
7. `DBI::dbDisconnect(con)`

`start_pipeline()` wires these together and handles per-batch errors so one failure does not stop the whole run.

## Data shapes

- **`yahoo_query_data()`** returns a *wide* tibble (one row per symbol × date, with `open`, `high`, …, `close_adjusted`, `index_ts`, `source`).
- **`format_data()`** pivots to *long* form with columns `index_ts`, `date`, `metric`, `value` (metrics: `open`, `high`, `low`, `close`, `volume`, `close_adjusted`).
- **`insert_new_data()`** skips rows whose `(date, index_ts, metric)` triple already exists in `data_sp500`.

## Further reading

- Function help: `?start_pipeline`, `?connect_db`  
- [README](https://github.com/pkinif/pipelineR) on GitHub  
- Issues: <https://github.com/pkinif/pipelineR/issues>
